Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-17 ICML2019/main.tex
move the proof into appendix
pass until section 3.3 included

\subsubsection{Using $\ell_1$ Norm with Improved Sensitivity}
 The accuracy of the previous $\lapmech$ can be improved, by adding noise scaled to smaller sensitivity.
 In terms of two adjacent data sets $\dataobs, \dataobs'$, the posterior distributions that they induce -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$. As a results, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.\\
 By calibrating noise to this improved sensitivity, $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces output:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)
$
with $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model and
$\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, 
\alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}
$
with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Mechanism \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

\subsection{Exponential Mechanism Family}
In this section, we explore the exponential mechanism family: $\expmech$ by considering Hellinger distance metrics and different sensitivities.\\
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$, we define the follwing set as candidate set where the mechanisms in this family sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs')\mid \dataobs' \in \{0,1\}^n\}.
\]
For simplicity:
$
  \candidateset\equiv\{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
$
over all $\dataobs \in \{0,1\}^n$, where $\Delta\alpha$ is as defined in Section \ref{sec_background}, and
$
  \candidateset\equiv\{\dirichlet{\valpha'} \mid \valpha'=\valpha + \Delta\valpha\}
$,
over all $\dataobs \in \mathcal{X}^n$ in Dirichlet-multinomial model.\\
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process when the other two parameters are fixed, denoted by $\bprior$ and $\binomial{n}{\theta}$, $\multinomial{n}{\vtheta}$.

>>>>

\subsection{Improving the sensitivity}
 The accuracy of $\lapmech$ can be reduced to a constant.
 Indeed, two adjacent data sets $\dataobs, \dataobs'$  induce posterior distributions -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ --
 whose parameters differ by at most 2 w.r.t $\ell_1$ norm. This because we can see the parameters of the distributions as an histogram.
 Hence, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.
 {\color{red} Notice that we are using the bounded notion of differential privacy \cite{????}.}
 \todo{why 1 in 2 dimensions?}
 We denote the mechanism which calibrates the noise to this smaller sensitivity as $\ilapmech$.
 $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0) $
 where $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in the Beta-binomial model and
 $\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots,  \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}$
 with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Figure \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

\subsection{A new mechanism}
In this section, we explore a new mechanism which we denote by $\expmech$.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs')\mid \dataobs' \in \{0,1\}^n\}.
\]
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.
